Ecological niche models of plant species exploited by Epipalaeolithic foragers in the \emph{Badia} (c.~14500 BP)

### 2023-04-26 analysis/analysis.tex
Updates from Feb 2020

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Ecological niche models of plant species exploited by Epipalaeolithic foragers in Badia (c.~14500 BP)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Ecological niche models of plant species exploited by Epipalaeolithic
foragers in \emph{Badia} (c.~14500 BP)}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"tidyverse"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Registered S3 methods overwritten by 'ggplot2':
##   method         from 
##   [.quosures     rlang
##   c.quosures     rlang
##   print.quosures rlang
\end{verbatim}

\begin{verbatim}
## -- Attaching packages ------------------------------------------------------------------------------------------ tidyverse 1.2.1 --
\end{verbatim}

\begin{verbatim}
## v ggplot2 3.1.1     v purrr   0.3.2
## v tibble  2.1.2     v dplyr   0.8.1
## v tidyr   0.8.3     v stringr 1.4.0
## v readr   1.3.1     v forcats 0.4.0
\end{verbatim}

\begin{verbatim}
## -- Conflicts --------------------------------------------------------------------------------------------- tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"magrittr"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Attaching package: 'magrittr'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:purrr':
## 
##     set_names
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:tidyr':
## 
##     extract
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"sf"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Linking to GEOS 3.7.1, GDAL 2.3.2, PROJ 5.2.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"raster"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: sp
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'raster'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:magrittr':
## 
##     extract
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:dplyr':
## 
##     select
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:tidyr':
## 
##     extract
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"broom"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"dismo"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"rnaturalearth"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"gt"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"BadiaPaleoFloraSDM"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sw_asia <-}\StringTok{ }\KeywordTok{extent}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{30}\NormalTok{, }\DecValTok{50}\NormalTok{, }\DecValTok{25}\NormalTok{, }\DecValTok{40}\NormalTok{))}
\NormalTok{s_levant <-}\StringTok{ }\KeywordTok{extent}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{34}\NormalTok{, }\DecValTok{39}\NormalTok{, }\DecValTok{29}\NormalTok{, }\DecValTok{34}\NormalTok{))}
\NormalTok{palestine <-}\StringTok{ }\KeywordTok{ne_countries}\NormalTok{(}\DataTypeTok{scale =} \DecValTok{10}\NormalTok{, }\DataTypeTok{country =} \KeywordTok{c}\NormalTok{(}\StringTok{"Israel"}\NormalTok{, }\StringTok{"Palestine"}\NormalTok{),}
             \DataTypeTok{returnclass =} \StringTok{"sf"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_union}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_sf}\NormalTok{()}

\NormalTok{land <-}\StringTok{ }\KeywordTok{read_sf}\NormalTok{(}\StringTok{"data/raw_data/ne/ne_10m_land.shp"}\NormalTok{)[,}\StringTok{"featurecla"}\NormalTok{] }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{st_crop}\NormalTok{(sw_asia)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersection assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning: attribute variables are assumed to be spatially constant
## throughout all geometries
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rivers <-}\StringTok{ }\KeywordTok{read_sf}\NormalTok{(}\StringTok{"data/raw_data/ne/ne_10m_rivers_lake_centerlines_scale_rank.shp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{st_crop}\NormalTok{(s_levant)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersection assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning: attribute variables are assumed to be spatially constant
## throughout all geometries
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{water <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{read_sf}\NormalTok{(}\StringTok{"data/raw_data/ne/ne_10m_ocean.shp"}\NormalTok{)[,}\StringTok{"featurecla"}\NormalTok{],}
               \KeywordTok{read_sf}\NormalTok{(}\StringTok{"data/raw_data/ne/ne_10m_lakes.shp"}\NormalTok{)[,}\StringTok{"featurecla"}\NormalTok{]) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{st_crop}\NormalTok{(sw_asia)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersection assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning: attribute variables are assumed to be spatially constant
## throughout all geometries
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{basemap <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ land, }\DataTypeTok{fill =}\StringTok{"grey"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ palestine, }\DataTypeTok{fill =} \StringTok{"lightgrey"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ rivers, }\DataTypeTok{colour =} \StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ water, }\DataTypeTok{fill =} \StringTok{"white"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_sf}\NormalTok{(}\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\KeywordTok{xmin}\NormalTok{(s_levant), }\KeywordTok{xmax}\NormalTok{(s_levant)),}
           \DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\KeywordTok{ymin}\NormalTok{(s_levant), }\KeywordTok{ymax}\NormalTok{(s_levant)))}
\NormalTok{basemap}
\end{Highlighting}
\end{Shaded}

\includegraphics{analysis_files/figure-latex/regions-1.pdf}

\hypertarget{occurrence-data}{%
\subsection{Occurrence data}\label{occurrence-data}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Download or read data from GBIF}
\NormalTok{gbif_path <-}\StringTok{ }\NormalTok{here}\OperatorTok{::}\KeywordTok{here}\NormalTok{(}\StringTok{"analysis/data/raw_data/gbif"}\NormalTok{)}
\NormalTok{occ <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Hordeum"}\NormalTok{, }\StringTok{"spontaneum*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Bolboschoenus"}\NormalTok{, }\StringTok{"glaucus*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),    }
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Bolboschoenus"}\NormalTok{, }\StringTok{"maritimus*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),    }
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Scirpus"}\NormalTok{, }\StringTok{"maritimus*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Triticum"}\NormalTok{, }\StringTok{"urartu*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Triticum"}\NormalTok{, }\StringTok{"dicoccoides*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Stipa"}\NormalTok{, }\StringTok{"*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Zilla"}\NormalTok{, }\StringTok{"spinosa*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Vitex"}\NormalTok{, }\StringTok{"*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path),}
  \KeywordTok{fetch_gbif}\NormalTok{(}\StringTok{"Fraxinus"}\NormalTok{, }\StringTok{"*"}\NormalTok{, }\DataTypeTok{ext =}\NormalTok{ sw_asia, gbif_path)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_hordeum_spontaneum.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_bolboschoenus_glaucus.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_bolboschoenus_maritimus.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_scirpus_maritimus.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_triticum_urartu.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_triticum_dicoccoides.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_stipa.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_zilla_spinosa.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_vitex.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{verbatim}
## Reading /home/joeroe/gdrive/Work/Lab/BadiaPaleoFloraSDM/analysis/data/raw_data/gbif/gbif_fraxinus.csv
\end{verbatim}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   gbifID = col_double(),
##   genus = col_character(),
##   species = col_character(),
##   lon = col_double(),
##   lat = col_double(),
##   basisOfRecord = col_character()
## )
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Code taxonomic groups for analysis}
\NormalTok{occ }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{taxon =} \KeywordTok{case_when}\NormalTok{(genus }\OperatorTok{==}\StringTok{ "Hordeum"} \OperatorTok{~}\StringTok{ "Hordeum spontaneum"}\NormalTok{,}
\NormalTok{                           genus }\OperatorTok{==}\StringTok{ "Bolboschoenus"} \OperatorTok{~}\StringTok{ "Bolboschoenus sp. (=Scirpus maritimus)"}\NormalTok{,}
\NormalTok{                           species }\OperatorTok{==}\StringTok{ "Triticum urartu"} \OperatorTok{~}\StringTok{ "Triticum urartu"}\NormalTok{,}
\NormalTok{                           species }\OperatorTok{==}\StringTok{ "Triticum dicoccoides"} \OperatorTok{~}\StringTok{ "Triticum dicoccoides"}\NormalTok{,}
\NormalTok{                           genus }\OperatorTok{==}\StringTok{ "Stipa"} \OperatorTok{~}\StringTok{ "Stipa sp."}\NormalTok{,}
\NormalTok{                           genus }\OperatorTok{==}\StringTok{ "Zilla"} \OperatorTok{~}\StringTok{ "Zilla spinosa"}\NormalTok{,}
\NormalTok{                           genus }\OperatorTok{==}\StringTok{ "Vitex"} \OperatorTok{~}\StringTok{ "Vitex sp."}\NormalTok{,}
\NormalTok{                           genus }\OperatorTok{==}\StringTok{ "Fraxinus"} \OperatorTok{~}\StringTok{ "Fraxinus sp."}\NormalTok{)) }\OperatorTok{%T>%}
\StringTok{  }\NormalTok{\{}\ControlFlowTok{if}\NormalTok{(}\KeywordTok{any}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(.}\OperatorTok{$}\NormalTok{taxon))) }\KeywordTok{stop}\NormalTok{(}\StringTok{"Taxon contains missing values!"}\NormalTok{)\} ->}
\StringTok{  }\NormalTok{occ}

\CommentTok{# Remove duplicate records within analytic groups}
\NormalTok{occ }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(taxon) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{distinct}\NormalTok{(lon, lat, }\DataTypeTok{.keep_all =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ungroup}\NormalTok{() ->}
\StringTok{  }\NormalTok{occ}

\CommentTok{# Convert to sf}
\NormalTok{occ }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{(}\DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\StringTok{"lon"}\NormalTok{, }\StringTok{"lat"}\NormalTok{), }\DataTypeTok{crs =} \DecValTok{4326}\NormalTok{, }\DataTypeTok{remove =} \OtherTok{FALSE}\NormalTok{) ->}
\StringTok{  }\NormalTok{occ}
\end{Highlighting}
\end{Shaded}

It's interesting to map the current distribution of these plants.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(occ, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ lon, }\DataTypeTok{y =}\NormalTok{ lat)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ occ, }\DataTypeTok{shape =} \DecValTok{1}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ water, }\DataTypeTok{inherit.aes =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{fill =} \StringTok{"white"}\NormalTok{, }\DataTypeTok{colour =} \StringTok{"darkgrey"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{taxon)}
\end{Highlighting}
\end{Shaded}

\includegraphics{analysis_files/figure-latex/occurrence-map-1.pdf}

\hypertarget{pseudo-absences}{%
\subsection{Pseudo-absences}\label{pseudo-absences}}

Occurrence data only tells us where a species is present; there is no
definitive information on where the species is \emph{not} found. We
therefore need to generate random background points or
``pseudo-absences'' to feed to the model. There are several ways to do
this. We follow the advice of Barbet-Massin et al.~(2012) for
regression-based species distribution models and simply use a large
(\textasciitilde10000) random sample of points.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{occ <-}\StringTok{ }\KeywordTok{as_tibble}\NormalTok{(occ)}
\NormalTok{occ}\OperatorTok{$}\NormalTok{occurrence <-}\StringTok{ }\OtherTok{TRUE}
\ControlFlowTok{for}\NormalTok{(taxon }\ControlFlowTok{in} \KeywordTok{unique}\NormalTok{(occ}\OperatorTok{$}\NormalTok{taxon)) \{}
\NormalTok{  occ <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(occ, }\KeywordTok{generate_psabsences}\NormalTok{(palestine, }\DataTypeTok{taxon =}\NormalTok{ taxon,}
                                            \DataTypeTok{occurrence =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{N =} \DecValTok{10000}\NormalTok{))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning in bind_rows_(x, .id): Vectorizing 'sfc_POINT' elements may not
## preserve their attributes

## Warning in bind_rows_(x, .id): Vectorizing 'sfc_POINT' elements may not
## preserve their attributes
\end{verbatim}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning in bind_rows_(x, .id): Vectorizing 'sfc_POINT' elements may not
## preserve their attributes
\end{verbatim}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning in bind_rows_(x, .id): Vectorizing 'sfc_POINT' elements may not
## preserve their attributes
\end{verbatim}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
\end{verbatim}

\begin{verbatim}
## Warning in bind_rows_(x, .id): Vectorizing 'sfc_POINT' elements may not
## preserve their attributes
\end{verbatim}

\begin{verbatim}
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
## although coordinates are longitude/latitude, st_intersects assumes that they are planar
\end{verbatim}

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Ecological niche models of plant species exploited by Epipalaeolithic foragers in the Badia (c.~14500 BP)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Ecological niche models of plant species exploited by Epipalaeolithic
foragers in the \emph{Badia} (c.~14500 BP)}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

\includegraphics{analysis_files/figure-latex/regions-map-1.pdf}

\hypertarget{background}{%
\subsection{Background}\label{background}}

\hypertarget{plant-exploitation-in-the-epipalaeolithic-southern-levant}{%
\subsubsection{Plant exploitation in the Epipalaeolithic Southern
Levant}\label{plant-exploitation-in-the-epipalaeolithic-southern-levant}}

\hypertarget{ecological-niche-modelling}{%
\subsubsection{Ecological niche
modelling}\label{ecological-niche-modelling}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

\hypertarget{occurrences}{%
\subsubsection{Occurrences}\label{occurrences}}

@@

It's interesting to map the current distribution of these plants.
\includegraphics{analysis_files/figure-latex/occurrence-map-1.pdf}

\hypertarget{pseudo-absences}{%
\subsubsection{Pseudo-absences}\label{pseudo-absences}}

Occurrence data only tells us where a species is present; there is no
definitive information on where the species is \emph{not} found. We
therefore need to generate random background points or
``pseudo-absences'' to feed to the model. There are several ways to do
this. We follow the advice of Barbet-Massin et al.~(2012) for
regression-based species distribution models and use a large
(\textasciitilde10000) random sample of points, weighted equally against
the presences in the regression.

\hypertarget{environmental-data}{%
\subsubsection{Environmental data}\label{environmental-data}}

\hypertarget{current-climate}{%
\paragraph{Current climate}\label{current-climate}}
